---
title: "WO1D"
author: "Benjamin Huang"
date: "2024-09-05"
output:
  pdf_document: default
  html_document: default
---

Standard Deviation is a measure of spread
$$
S_Y = \sqrt{\frac{1}{n-1}\sum(y_i-\bar{y})^2}
$$

```{r}
#assigns values to A
A <- c(4, 5, 6)
#sample deviation of A
sd(A)
```

Q1. Use the code above to find the standard deviation of B=(3, 5, 7)
```{r}
B <- c(3, 5, 7)
sd(B)
```

Correlation describes the linear relationship between A and B.
$$
r_{X,Y} = \frac{1}{(n-1)s_Xs_Y}\sum(x_i-\bar{x})(y_i-\bar{y})
$$

Q2. Use the cor() function to find the correlation between A and B
```{r}
A <- c(4, 5, 6)
B <- c(3, 5, 7)
cor(A, B)
```

Sometimes you need to use a function that isn't in base R. Fortunately, R is a open resource and many packages have been created to solve these problems.
```{r}
#install.packages("moments")
library(moments)
A <- c(4, 5, 6)
skewness(A)
```

A is symmetrical around its mean, 5.

Q3. Plot and find the skewness of the data set pressure. Is this data normally distributed?
```{r}
data("pressure")
plot(pressure)
```

A histogram can make it clear if a data set is normally distributed.
```{r}
set.seed(441)
par(mfrow = c(1,2))
hist(pressure$pressure)
hist(rnorm(1000))
#they do not look similar
```

Another graphical way of checking if data is normally distributed is with a qq plot (Quantile-Quantile plot). The idea is to plot the percentiles, or quantiles, of one distribution against the other. If they come from the same distribution, then the points should lie on a line
```{r}
qqnorm(pressure$pressure); qqline(pressure$pressure)
#not normally distributed
```

Q4. Below is a block of code that shows what a qq plot will look like for normally distributed data. Describbe what each line does in front of the comments, #.
```{r}
#
set.seed(441)
n <- 50
(x <- rnorm(n, 0, 1))
qqnorm(x); qqline(x)
```

The pnorm(x) function gives $P(Z<x)$ for a normal distribution.
The qnorm(p) function gives the x-value for a given percentile.
```{r}
pnorm(1.96)
qnorm(0.975)
```

Q5. Find the probability that a draw from a standard normal distribution is between -1.5 and 0.5.
```{r}
pnorm(0.5) - pnorm(-1,5)
```

Q6. Determine the z-values that contain the middle 60% of the standard normal distribution.
```{r}
qnorm(0.2)
qnorm(0.8)
```




