---
title: "W02D"
author: "Benjamin Huang"
date: "2024-09-05"
output:
  pdf_document: default
  html_document: default
---

The coefficients in a basic linear regression can be calculated as

\begin{align*}
& b_1 = r\frac{s_y}{s_x},
\\
& b_0 = \bar{y} - b_1\bar{x}
\end{align*}

Below is an example data set and a calculation of the parameters for B~A
```{r}
set.seed(441)
A <- 1:10
e <-  rnorm(10, 0, 1)
B <- 2*A - 1 + e
(b_1 <- cor(A, B)*sd(B)/ sd(A))
(b_0 <- mean(B)-b_1*mean(A))
```

Q1. What are the population parameters $\beta_0$ and $\beta_0$ ? Why don't they equal the sample parameters $b_0$ and $b_1$ ?

Answer:

\begin{align*}
& Y_1 = \beta_0 + \beta_1 X_1
\\
& Y = b_0 + b_1X + \epsilon
\\
& \epsilon \backsim N(0,1)
\end{align*}

Below, a regression is run to regress B on A. Then a summary is produced.
```{r}
lm <- lm(B~A)
summary(lm)
```

Q2. What does "lm" stand for?

Answer: linear model.

The residual standard error is $s = 1.198$, our residual standard deviation. Note there are $n-2=8$ degrees of freedom for $s$ .
The Multiple R-squared is $R^2 = 0.952$ , the coefficient of determination.

Q3. What is an estimate for the variance of disturbances? What proportion of the variance of B is explained by A?

Answer:

\begin{align*}
& Estimate \: of \: disturbances: e_i = Y_i - \hat{Y_i}
\\
& S_e = \frac{\sum(y_i-\hat{y_i})^2}{n-2}
\\
& \sigma^2 \approx s^2 = 1.43.
\\
& R^2 = 0.952
\end{align*}

Below is an example of an Analysis of Variance (ANOVA) table.  Note that the mean square error is 1.43.
```{r}
anova(lm)
```

Q4. Compute the total sum of squares. Relate this to $R^2$ the variance of B.
```{r}
(TSS <- 227.405 + 11.476)
(r_square <- 227.405 / TSS)
```

Below is a plot of the points along with the best fit line.
```{r}
plot(A, B) # plots the points
abline(a=coef(lm)[1], b=coef(lm)[2], col = "green")
#abline(lm, col = "blue")
```

Doubling the Explanatory Variable

Q5. Predict which of the following values will change when we double the values of A: $S_A, s_B, r_{AB}, b_0, b_1, s, R^2$

Answer:

```{r}
#Answer:
A2  <- 2*A
lm2 <- lm(B~A2)
summary(lm2)
anova(lm2)
```

Mathematical Explanations:
If A is the only thing that changed, then $s_A$ will change by the same factor.
The correlation will stay the same because the sum is doubled and $s_A$ double, canceling any change.

$r = \frac{1}{(n-1)s_As_B}\sum(A_i-\bar{A})(B_i-\bar{B}) = \frac{1}{(n-1)(2*s_A)s_B}\sum(2*A_i-s*\bar{A})(B_i- \bar{B})$

We have $b_1 = r\frac{s_B}{s_A} \rightarrow r\frac{s_B}{2*s_A}$, so $b_1$ is halved.

$b_0$ stays the same because the *2 in $\bar{x}$ cancels with the 1/2 in $b_1$.

$b_0 = \bar{y} - b_1*\bar{x} = \bar{y} - \frac{1}{2}b_1*(2*\bar{x})$

Below is the plot of (A2, B) and the best fit line.
```{r}
plot(A2, B) # plots the points
abline(a=coef(lm2)[1], b=coef(lm2)[2])
```

Doubling the Variable of Interest

Q6. Predict which of the following values will change when we double the values of B: $S_A, s_B, r_{AB}, b_0, b_1, s, R^2$

Answer:

```{r}
#Answer:
B2 <- 2*B
lm3 <- lm(B2~A)
summary(lm3)
anova(lm3)
```

Below is a plot of (A, B2) and the best fit line.
```{r}
plot(A, B2) # plots the points
abline(a=coef(lm3)[1], b=coef(lm3)[2])
```

Double Both

Q6. Predict which of the following values will change when we double the values of A and B

Answer:

```{r}
#Answer:
lm4 <- lm(B2 ~ A2)
summary(lm4)
anova(lm4)
```

```{r}
plot(A2, B2) # plots the points
abline(a=coef(lm4)[1], b=coef(lm4)[2])
```
